Scalable Clustering of Genotype Information using MapReduce

### 2014-05-14 doc/publications/incob2014/paper.tex
Updates to methods

\section*{Methods}
\subsection*{Datasets}
Trio data from the 1000 genomes project -> determine in the clustering can identify ancestry 
Sequence data from hapmap samples -> determine ethnic groups
Scaled down version of the hap map data (Chr1) for R to run on
Overview see Table~\ref{datasets}

\subsection*{K-means clustering}
We Cluster samples from a VCF file containing the variants from chromosome one. The content of the VCF file represents
Mahout provides libraries for clustering data, 'sequences', although the data must first be converted from the VCF format to
sequence files. The sequence files should consist of key-value pairs, where
Using a custom Java program, we transpose the

>>>>

\section*{Methods}
\subsection*{Datasets}
For k-means clustering of individuals, we used variant call format (VCF) files from the 1000 Genomes Project.
These VCF files contain information about the genetic variants between each of the 1092 individuals and a reference genome
derived from GRCh37. Each of the VCF files are partitioned by chromosome.
We used the VCF file containing chromosome 1 variants (3,007,196 variants) as a subset of the entire human variant dataset.
For the entire dataset, we used each of the chromosomal VCF files. Metadata available from the 1000 Genomes Consortium
specifies additional attributes for each individual, including population and familial data.
Table~\ref{datasets} contains an overview of the datasets.

> determine in the clustering can identify ancestry 
Sequence data from hapmap samples -> determine ethnic groups
Overview see Table~\ref{datasets}

\subsection*{Converting to arrays}
Mahout requires sequence files as input to its clustering algorithms and provides libraries for writing vectors to these files.
These libraries require a feature vector and optionally a sample name. The vectors must contain an integer for each element.
Therefore, we need to preprocess the data from the VCF files into a suitable format. The variant data values, which are tab separated,
must be converted to an integer. For this, we simply determine the Hamming distance between each tab separated value, and the reference genome.
Therefore, a value identical to the reference will therefore be 0, a heterozygous variant will be 1 and a homozygous variant will be 2.

\subsection*{Mahout K-means clustering}
We Cluster samples from a VCF file containing the variants from chromosome one. The content of the VCF file represents

Mahout provides libraries for clustering data, 'sequences', although the data must first be converted from the VCF format to
sequence files. The sequence files should consist of key-value pairs, where
Using a custom Java program, we transpose the

\subsection*{Clustering quality}
We scored the clusters with the Adjusted Rand index. This index compares two different clusterings and assigns a value based
on their similarity. The value is zero for independent clusterings and one for identical clusterings. We compared the mahout k-means
clusters and the R k-means clusters to the known clusters based on meta data from the 1000 genome project. For the known clusters,
individuals were partitioned by their super population code (ASN, EUR, AFR, AMR). We also compared the mahout clusters to the
R clusters for the Chr1 subset data.

\subsection*{Speed}
